---
title: "ECON457 lab02 R For Economics"
author: "Jon"
date: "September 23, 2016"
output: html_document
---


### A series of video of R tutorial. 


[MarinStatsLectures](https://www.youtube.com/user/marinstatlectures/videos)


## R for ECON

### R for ECON345

```{r}
library(foreign)
wage1 <- read.dta("http://fmwww.bc.edu/ec-p/data/wooldridge/wage1.dta")

# OLS regression:
ols.model <-  lm(log(wage) ~ educ+exper+tenure, data=wage1) 

```

```{r}
# Different functions to work with the model
# Print a summary of the results
summary(ols.model)

# Produce confidence intervals
confint(ols.model)

# Get a vector of the coefficients
coef(ols.model)
```


### R for ECON345/365/366

Matrix transposition

```{r}
M <- cbind(c(1,0,1),c(0,1,2),c(0,0,1))
t(M)

```

Matrix inversion

```{r}
solve(M)

solve(M)%*%M
```



### R for ECON208

Solving a linear equation
```{r}
m=matrix(nrow=2,ncol=2,c(1,-.8,1,.2))

l=matrix(c(1.0+25.0/18,25.0/18.0))

(k=solve(m,l))

m%*%k          #checking the answer


```

Eigenvalues and eigenvectors

```{r}
eigen(m)
```



### R for ECON350/351

Symbolic calculations

```{r}
D(expression(x^n),"x")
D(expression(exp(a*x)),"x")
```


Compute derivatives of simple expressions, symbolically and algorithmically.


```{r}
trig.exp <- expression(sin(cos(x + y^2)))

( D.sc <- D(trig.exp, "x") )

all.equal(D(trig.exp[[1]], "x"), D.sc)
```



```{r}
## formula argument :
(MU1 <- deriv(~ x1^0.25*x2^0.25, "x1" )) ; 
(MU2<- deriv(~ x1^0.25*x2^0.25, "x1" ))

mode(MU1)
x1<- 2
x2 <- 2
eval(MU1)



```



#### Unconstrained optimization

```{r}
f <- function(x){(x[1] - 5)^2 + (x[2] - 6)^2}
initial_x <- c(10, 11)
x_optimal <- optim(initial_x, f, method="CG")
x_min <- x_optimal$par
x_min
```



## ECON457

#### Equality constraint optimization
#### Inequality constraint optimization
#### Linear Programming


Maximize expected return: f(x1, x2, x3) = x1*5% + x2*4% + x3*6%

Subjected to constraints:
    10% < x1, x2, x3 < 100%
    x1 + x2 + x3 = 1
    x3 < x1 + x2
    x1 < 2 * x2

```{r}
#install.packages("lpSolve")
#install.packages("lpSolveAPI")
library(lpSolve)
library(lpSolveAPI)
# Set the number of vars
model <- make.lp(0, 3)
# Define the object function: for Minimize, use -ve
set.objfn(model, c(-0.05, -0.04, -0.06))
# Add the constraints
add.constraint(model, c(1, 1, 1), "=", 1)
add.constraint(model, c(1, 1, -1), ">", 0)
add.constraint(model, c(1, -2, 0), "<", 0)
# Set the upper and lower bounds
set.bounds(model, lower=c(0.1, 0.1, 0.1), upper=c(1, 1, 1))
# Compute the optimized model
solve(model)

# Get the value of the optimized parameters
get.variables(model)

# Get the value of the objective function
get.objective(model)

# Get the value of the constraint
get.constraints(model)
```

#### Quadratic Programming


To formulate the problem:

Variable: x1 = % investment in A, x2 = % in B, x3 = % in C

Minimize variance: x

Subjected to constraint:
x1 + x2 + x3 == 1
X1*5% + x2*4% + x3*6% >= 5.2%
0% < x1, x2, x3 < 100%

```{r}
#install.packages("quadprog")
library(quadprog)
mu_return_vector <- c(0.05, 0.04, 0.06) 
sigma <- matrix(c(0.01, 0.002, 0.005, 
                   0.002, 0.008, 0.006, 
                   0.005, 0.006, 0.012), 
                  nrow=3, ncol=3)
D.Matrix <- 2*sigma
d.Vector <- rep(0, 3)
A.Equality <- matrix(c(1,1,1), ncol=1)
A.Matrix <- cbind(A.Equality, mu_return_vector, 
                    diag(3))
b.Vector <- c(1, 0.052, rep(0, 3))
out <- solve.QP(Dmat=D.Matrix, dvec=d.Vector, 
                  Amat=A.Matrix, bvec=b.Vector, 
                  meq=1)
out$solution

out$value

```


[source][4]

### Ryacas: Symbolic calculation

only works for R 3.3.2

```{r}
# # options(repos = c(CRAN = "http://cran.rstudio.com"))
# # install.packages('Ryacas')
# library(Ryacas)
# 
# #package?Ryacas
# print(yacas(expression(integrate(1/x, x))))
# print(yacas("Integrate(x) 1/x"))
# x <- Sym("x"); Integrate(1/x, x)
# acos(Sym("1/2"))
# my_func <- function(x) {
#     return(x/(x^2 + 3))
# }
# my_deriv <- yacas(deriv(my_func(x), x))
```



### R for ECON435/468

http://www.quantmod.com/
https://www.youtube.com/watch?v=61_F2fcvrsw

```{r}
#install.packages("quantmod")
library(quantmod)
getSymbols(c("AMZN"))
barChart(AMZN,theme='white.mono',bar.type='hlc')

```



```{r}
getSymbols(c("^GSPC"))
chartSeries(GSPC, subset='last 3 months')
addBBands(n = 20, sd = 2, ma = "SMA", draw = 'bands', on = -1)
```

See what packages we used.

```{r}
sessionInfo()
```


[1]:http://www.levelbased.com/guides/pokemon-go/pokemon/?q=
[2]: http://www.levelbased.com/static/images/guides/pokemon-go/1/medium/bulbasaur.png "Bulbasaur"
[3]: https://github.com/Eighty20/eighty20.github.io/raw/master/_rmd/Post_data/All_pokemon.csv "Pokeman.csv" 
[4]: http://horicky.blogspot.ca/2013/01/optimization-in-r.html