---
title: "ECON457 lab01 R Basic"
author: "Jon"
date: "September 16, 2016"
output: html_document
#output: statsr:::statswithr_lab

---

# Basic R

Best to start a new project. 


## Introduction to R Studio

## Mathematical Operations

Below is a list of common mathematical operations that you can perform on numerical types. 


x + y		performs addition


x - y		performs subtraction 


x * y		performs multiplication 


x / y		performs division

x ^ y		raises x to the yth power



x = y		assigns the variable named "x" to the value y


x == y	evaluates to a Boolean, true if x equals y, false otherwise

x != y 	evaluates to a Boolean, true if x does not equal y, false otherwise


x > y		evaluates to a Boolean, true if x is greater than y, false otherwise


x < y		evaluates to a Boolean, true if x is less than y, false otherwise


x <= y	evaluates to a Boolean, true if x is less than or equal to y, false otherwise


x >= y	evaluates to a Boolean, true if x is greater than or equal to y, false otherwise

```{r}
9-4
9 / 4
```


```{r}
log(exp(10))
```



```{r}
sin(pi/2)
```




## Data type 

R has five basic or "atomic" classes of objects:

* character

* numeric (real numbers)

* integer

* complex

* logical (True/False)


```{r}
x <- c(0.5, 0.6);  typeof(x)       ## numeric
x <- c(TRUE, FALSE);  typeof(x)    ## logical
x <- c(T, F);   typeof(x)         ## logical
x <- c("a", "b", "c");  typeof(x)   ## character
x <- c(1+0i, 2+4i);  typeof(x)    ## complex
x <- 9:29;   typeof(x)            ## integer
```


`":"` operator can create a sequence which is very useful.


`seq()` can do similar thing and is more powerful.

```{r}
seq(9,29, by=2)
```



## Data Structure


### Vector


The most basic type of R object is a vector. Empty vectors can be created with the vector() function or c() function. There is really only one rule about vectors in R, which is that A **vector can only contain objects of the same class**.

You can also use the vector() function to initialize vectors.


```{r}
x <- vector("numeric", length = 10) 
x
```


Question 1: Generate a vector 

$$a =  (3, 6, 9 ,12) $$

```{r q1}
# type your code for Question 1 here, and Knit




```





**A vector can only contain objects of the same class.**



**A matrix can only contain objects of the same class.**




### Matrix

Initial a empty matrix

```{r}
m <- matrix(nrow = 2, ncol = 3) 
m
```

**vector -> matrix**

Matrices can be created by column-binding or row-binding with the cbind() and rbind() functions.


```{r}
x <- 1:3
y <- 10:12
cbind(x, y)
```



```{r}
rbind(x, y) 
```


Question 2: Generate a matrix  

\begin{matrix}
  1 & 2 & 3 \\
  4 & 5 & 6 \\
  7 & 8 & 9
\end{matrix}

```{r q2}
# type your code for Question 1 here, and Knit




```






### Data Frames


Data frames can contain objects of different classes.

Data frames are used to store tabular data in R. They are an important type of object in R and are used in a variety of statistical modeling applications. 



```{r}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F)) 
x
```



## Check what is inside an object


```{r}
length(y)
dim(y)
str(y)
attributes(y)
typeof(y)
head(y, 1)
tail(y,1)
```


More for data frame.

```{r}
nrow(x)

ncol(x)

names(x)
```



Using `str()` function to check `x`.

```{r}
str(x)
```

* What class is the `foo` column?

* What class is the `bar` column?


** Different classes of data can be in a data frame**



## Load Data from Csv File

R works with many data formats. 

csv file is the most convenient.


### Load data locally 


Download file from [here][3]


```{r}
if (!file.exists("All_pokemon.csv")){
   download.file("https://github.com/Eighty20/eighty20.github.io/raw/master/_rmd/Post_data/All_pokemon.csv", destfile = "pokemon.csv",method="wininet")
}

```



```{r}
pokemon <- read.csv("All_pokemon.csv")
```



Check data

```{r}
head(pokemon)
```


```{r}
str(pokemon)
```



###  directly from online repository.



```{r}
library(readr) # prefered
pokemon <- read_csv("https://github.com/Eighty20/eighty20.github.io/raw/master/_rmd/Post_data/All_pokemon.csv")
```


### Write 


More information about [pokeman][1] 

![Bulbasaur][2]



### Subsetting


#### Indexing


Indexing is used to specify the elements of an array. Indexing also allows you to get out certain bits of information from an array. To index into an array, type the name of the array, followed by the index of the element you want in brackets. Note that in R, indices start at 1. 

For a multidimensional array, index by [row,column]

To index an entire row or column, use a colon.

Below we index into the named y to get out the element in the second row, third column, 6.


```{r}
y <- c( 1,2,3, 4,5,6)
y <- matrix(y, nrow = 2, ncol = 3, byrow = T)
y[2,3]
```



If you want to change the value in y[2,3], you can assign a new value to it.


```{r}
y[2,3] <- 7
y
```



Below, we index the entire first row of the array named y.


```{r}
y[1,]
```


Below, we index the entire second column of the array named y.


```{r}
y[,2]
```


#### Subset   by name,
    
For data frame, `$` to subset

```{r}
type2 <- pokemon$Type.II
head(type2)
```


Look those levels

```{r}
table(type2)
```





#### Subset by logic indexing or comparison operations

```{r}
Dragon <-  pokemon[ pokemon$Type.I == "Dragon", ]

head(Dragon)

```

Question 3: Subset pokeman with HP greater than 150  


```{r  q3}
# type your code for Question 1 here, and Knit


```



### Write data


```{r}
#if (!file.exists("Dragon.csv")){
write.csv(Dragon, file="Dragon.csv")
#}
```


By default R will wrap character vectors with quotation marks when writing out to file. It will also write out the row and column names.

Let's fix this:


```{r}
#if (!file.exists("Dragon.csv")){
write.csv(Dragon,file="Dragon.csv",
          quote=FALSE, row.names=FALSE)
#}
```



## Control Flow

### Condition  if/elseif/else 

Conditional evaluation allows portions of code to be evaluated or not evaluated depending on the value of a Boolean expression. You do not need all if/elseif/else statements. You can have conditional evaluations with just an if, or just an if/else.

The general structure of conditional evaluation is as follows.

After assigning values to x and y and running the code we obtain the following output.

```{r}
x =1
y=2
if (x<y){
  print(x)
}else{
  print(y)
}


```




Question 4: Write a if else statement to  print larger number in x and y. 


```{r  q4}
# type your code for Question 1 here, and Knit


```




### For Loops 

A for loop allows you to specify the number of iterations for the repeated execution of a code block. They are great when you know how many iterations you want to run. 

The general form of a for loop is shown below. The example shows a for loop that calculates the sum of the integers 1 through 10 and prints the final result. 

Note that to obtain a range of integers, we use the colon : symbol. 

```{r}
sum = 0
for(num in 1:10){
  sum = sum + num
}
print(sum)
```



Question 5: Write a for loop statement to print 1 to 10


```{r  q5}
# type your code for Question 1 here, and Knit


```




## Plot 

```{r}
x = seq(0, 2*pi, 2*pi/180)
y = sin(x)
plot(x,y)

```


#### Color, line type, line width and title
```{r}
plot(x,y, type = "l", col="blue", lwd="4",
     main="sin", xlab = "x", ylab = "y")
```

### Add two lines and a point.


```{r}
plot(x,y, type = "l", col="blue", lwd="4",
     main="sin", xlab = "x", ylab = "y")
## the x- and y-axis, and an integer grid
abline(h = 0, v = pi, col = "gray60")
points(x=pi, y=0, cex =2, col ="dark red")
text(x=pi+0.3, y = 0.1, labels = "Here", col = "Red")
```



Question 6: Plot pokemon's HP and Atk with green color.
Change label and title.


```{r  q6}
# type your code for Question 1 here, and Knit


```




### How to get help


```{r}
?plot
```






### Learn by doing

Visit [DataCamp][4] or [Try R][5] to learn how to write basic R code. Both sites provide interactive lessons that will get you writing real code in minutes. They are a great place to make mistakes and test out new skills. You are told immediately when you go wrong and given a chance to fix your code.

### A series of video of R tutorial. 

Each tutorial usually only is 2 minutes long.

http://www.twotorials.com/


See what packages we used.

```{r}
sessionInfo()
```


[1]:http://www.levelbased.com/guides/pokemon-go/pokemon/?q=
[2]: http://www.levelbased.com/static/images/guides/pokemon-go/1/medium/bulbasaur.png "Bulbasaur"
[3]: https://github.com/Eighty20/eighty20.github.io/raw/master/_rmd/Post_data/All_pokemon.csv "Pokeman.csv" 
[4]:https://www.datacamp.com/
[5]:http://tryr.codeschool.com/
