---
title: "ECON457 R lab 3 Optimization in R"
output:
  html_notebook: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Information about optimizaion in R

http://cran.r-project.org/web/views/Optimization.html

https://www.rdocumentation.org/packages/ROI/versions/0.2-1


## The R Optimization Infrastructure (ROI)

The R Optimization Infrastructure (ROI) package promotes the development and use of interoperable (open source) optimization problem solvers for R.


ROI handle LP up to MILP and MIQCP problems using the following supported solvers :

1. lpSolve

2. quadprog

3. Rcplex

4. Rglpk (default)

5. Rsymphony

more... 




## Solvers

CPLEX is a commercial solver. Students can get academic license. 

GLPK is free. R has the high level interface Rglpk.

http://winglpk.sourceforge.net/

https://cran.r-project.org/web/packages/Rglpk/index.html



## Install all the packages.

```{r}
#install.packages("ROI")
#install.packages(c( "Rglpk","ROI.plugin.symphony","ROI.plugin.glpk","ROI.plugin.quadprog","ROI.plugin.nloptr","ROI.plugin.ipop","ROI.plugin.ecos"))
```




## Linear Progarmming





```{r}


library("ROI")

#library("ROI.plugin.glpk")
#library("ROI.plugin.symphony")
#library(ROI.plugin.ipop)

#library(ROI.plugin.nloptr)
#library(ROI.plugin.quadprog)
#library(ROI.plugin.ecos)

ROI_registered_solvers()
```



$$\begin{align*} 
    \text{maximize: }   2 x_1 &+ 4 x_2 + 3 x_3\\
\text{subject to: }     3 x_1 &+ 4 x_2 + 2 x_3 & <= 60 \\
                        2 x_1 &+   x_2 +   x_3 & <= 40 \\
                          x_1 &+ 3 x_2 + 2 x_3 & <= 80

\end{align*}$$




```{r}
## Simple linear program.
## maximize:   2 x_1 + 4 x_2 + 3 x_3
## subject to: 3 x_1 + 4 x_2 + 2 x_3 <= 60
##             2 x_1 +   x_2 +   x_3 <= 40
##               x_1 + 3 x_2 + 2 x_3 <= 80
##               x_1, x_2, x_3 are non-negative real numbers

LP <- OP( c(2, 4, 3),
          L_constraint(L = matrix(c(3, 2, 1, 4, 1, 3, 2, 2, 2), nrow = 3), # linear constraint
                       dir = c("<=", "<=", "<="), # direction
                       rhs = c(60, 40, 80)),      # right hand side
          max = TRUE )
LP
```

#### Solve

```{r}
sol <- ROI_solve(LP)# , solver = "glpk")
sol
```

#### Solutions

```{r}
solution(sol, type = c("primal"))
solution(sol, type = c("dual"))
```


#### Extract solutions
```{r}
sol$solution
sol$objval
sol$status
sol$message
```

## Quadratic Progarmming


It will require putting problem in matrix form.



$$\begin{align*} 
    \text{minimize: }   - 5 x_2 + 1/2 (x_1^2 + x_2^2 + x_3^2)\\
\text{subject to: }     -4 x_1 - 3 x_2       & >= -8 \\
                         2 x_1 +   x_2       & >= 2 \\
                         - 2 x_2 + x_3 & >= 0

\end{align*}$$



```{r}
## Simple quadratic program.
## minimize: - 5 x_2 + 1/2 (x_1^2 + x_2^2 + x_3^2)
## subject to: -4 x_1 - 3 x_2       >= -8
##              2 x_1 +   x_2       >=  2
##                    - 2 x_2 + x_3 >=  0

QP <- OP( Q_objective (Q = diag(1, 3), L = c(0, -5, 0)),   # quadratic objective
          L_constraint(L = matrix(c(-4,-3,0,2,1,0,0,-2,1), #  linear constraint
                                  ncol = 3, byrow = TRUE), 
                       dir = rep(">=", 3), 
                       rhs = c(-8,2,0)) )
QP
```

#### Solve

```{r}
sol <- ROI_solve(QP, solver = "quadprog")
sol
```


#### Extract solutions
```{r}
sol$solution
sol$objval
sol$status
sol$message
```


#### Solutions

```{r}
solution(sol, type = c("primal"))
solution(sol, type = c("dual"))
```



## Portfolio optimization - minimum variance

```{r}
## Portfolio optimization - minimum variance
## -----------------------------------------
## get monthly returns of 30 US stocks
data( US30 )
r <- na.omit( US30 )
## objective function to minimize
obj <- Q_objective( 2*cov(r) ) # quadratic objective
## full investment constraint
full_invest <- L_constraint( rep(1, ncol(US30)), "==", 1 ) #  linear constraint
## create optimization problem / long-only

(op <- OP( objective = obj, constraints = full_invest )) 
```

#### Solution


```{r}

## solve the problem - only works if a QP solver is registered
##  
res <- ROI_solve( op )
res
sol <- solution( res )
names( sol ) <- colnames( US30 )
round( sol[ which(sol > 1/10^6) ], 3 )
```




### LP with boundary


```{r}
suppressMessages(library(ROI))
lp_obj <- L_objective(c(1, 2)) #  linear objective
 lp_con <- L_constraint(c(1, 1), dir="==", rhs=2)  #  linear constraint
 lp_bound <- V_bound(ui=1:2, ub=c(3, 3))
 #ui an integer vector specifying the indices of non-standard (i.e., values != Inf) upper bounds.
 # ub up bound
 lp <- OP(objective=lp_obj, constraints=lp_con, bounds=lp_bound, maximum=FALSE)
 bounds(lp)
lp
```

 
#### Solutions 
 
```{r}
 x <- ROI_solve(lp)
 x$objval
 x$solution
```

####  Change boundary 

```{r}
# change boundary 
 bounds(lp) <- V_bound(ui=1:2, ub=c(1, 1))

```


#### Solutions 

```{r}
 y <- ROI_solve(lp)
 y$objval
 y$solution
```


--------------------------------------------------------------------------------------------

## Other optimization package in R


We also can solve optimization problem using other packages.

### Install packages

```{r}
#install.packages(c("lpSolve", "lpSolveAPI"))
#install.packages(c("quadprog","Rglpk", "nlme"))
```



```{r}

suppressMessages(library(lpSolve))

suppressMessages(library(lpSolveAPI))
suppressMessages(library(quadprog))
#suppressMessages(library(nlme))

suppressMessages(library(Rglpk))
```

```{r}
args(lp)
args(solve.QP)
args(Rglpk_solve_LP)
#args(nlminb)
#args(optim)
#args(Rcplex)
```


###  Linear Programming by lpsolve

Here is a list of some key features of lp_solve:

- Mixed Integer Linear Programming (MILP) solver

- Basically no limit on model size

- It is free and with sources

- Supports Integer variables, Semi-continuous variables and Special Ordered Sets


#### Example by lpSolveAPI

straightforward



$$\begin{align*} 
    \text{minimize: }   x_1 + 3 x_2 + 6.24 x_3 + 0.1 x_4\\
    \text{subject to: }     78.26 x_2 + 2.9 x_4       & >= 92.3 \\
                         0.2 x_1 +  11.91 x_3       & <= 14.8 \\
                         12.68 x_1 + 0.08 x_3 + 0.9 x_4 &>= 4 \\
                        18  <=  x_4 & <= 48.98 \\
                          x_1 & >=  28.6 
\end{align*}$$


##### Set up model

```{r}
library(lpSolveAPI)
#create an empty model
lpmodel <- make.lp(nrow =0, ncol= 4)
# nrow	: # a nonnegative integer value specifying the number of constaints in the linear program.
# ncol	: # a nonnegative integer value specifying the number of decision variables in the linear program.


```

##### Setup objective function

```{r}
# object
set.objfn(lpmodel, c(1, 3, 6.24, 0.1))
```

##### Add constraints

```{r}

#constraint
add.constraint(lpmodel, c(0, 78.26, 0, 2.9), ">=", 92.3)
add.constraint(lpmodel, c(0.24, 0, 11.31, 0), "<=", 14.8)
add.constraint(lpmodel, c(12.68, 0, 0.08, 0.9), ">=", 4)

```


##### Set bounds

```{r}
set.bounds(lpmodel, lower = c(28.6, 18), columns = c(1, 4))
set.bounds(lpmodel, upper = 48.98, columns = 4)

```


##### Setup names 


```{r}
RowNames <- c("THISROW", "THATROW", "LASTROW")
ColNames <- c("COLONE", "COLTWO", "COLTHREE", "COLFOUR")
dimnames(lpmodel) <- list(RowNames, ColNames)
lpmodel
```




##### Solve the model

```{r}
solve(lpmodel)

```



##### Extract solutions


```{r}
get.objective(lpmodel)
  # [1] 31.78276
  # 
get.variables(lpmodel)
  # [1] 28.60000  0.00000  0.00000 31.82759
  # 
get.constraints(lpmodel)
  # [1]  92.3000   6.8640 391.2928
```



#### Example by lpSolve

provide a lot of information including sensitivities analysis. 

$$\begin{align*} 
    \text{maximize: }   x_1 + 9 x_2 +  x_3 \\
    \text{subject to: }     x_1 + 2 x_2 + 3 x_3       & <= 9 \\
                         3 x_1 + 2 x_2  + 2 x_3       & <= 15  
\end{align*}$$


##### Setup objective function

```{r}
library(lpSolve)
#
# Set up problem: maximize
#   x1 + 9 x2 +   x3 subject to
#   x1 + 2 x2 + 3 x3  <= 9
# 3 x1 + 2 x2 + 2 x3 <= 15
#
model_obj <- c(1, 9, 1)
```

##### Setup constraints

```{r}
model_con <- matrix (c(1, 2, 3, 3, 2, 2), nrow=2, byrow=TRUE)
model_dir <- c("<=", "<=")
model_rhs <- c(9, 15)

```

##### Solve the model

```{r}
#
lp ("max", model_obj, model_con, model_dir, model_rhs)
##  Success: the objective function is 40.5

```


##### Extract solutions

```{r}

lp ("max", model_obj, model_con, model_dir, model_rhs)$solution
```

######  Get sensitivities

```{r}


# Get sensitivities
#
lp ("max", model_obj, model_con, model_dir, model_rhs, compute.sens=TRUE)$sens.coef.from
##  [1] -1e+30  2e+00 -1e+30
lp ("max", model_obj, model_con, model_dir, model_rhs, compute.sens=TRUE)$sens.coef.to  
##  [1] 4.50e+00 1.00e+30 1.35e+01
#
# Right now the dual values for the constraints and the variables are
# combined, constraints coming first. So in this example...
#
lp ("max", model_obj, model_con, model_dir, model_rhs, compute.sens=TRUE)$duals     
##  [1]   4.5   0.0  -3.5   0.0 -10.5
#
# ...the duals of the constraints are 4.5 and 0, and of the variables,
# -3.5, 0.0, -10.5. Here are the lower and upper limits on these:
#
lp ("max", model_obj, model_con, model_dir, model_rhs, compute.sens=TRUE)$duals.from
##  [1]  0e+00 -1e+30 -1e+30 -1e+30 -6e+00
lp ("max", model_obj, model_con, model_dir, model_rhs, compute.sens=TRUE)$duals.to  
##  [1] 1.5e+01 1.0e+30 3.0e+00 1.0e+30 3.0e+00
#


```

#### Intergr Programming

```{r}
# Run again, this time requiring that all three variables be integer
#
lp ("max", model_obj, model_con, model_dir, model_rhs, int.vec=1:3)
##  Success: the objective function is 37
lp ("max", model_obj, model_con, model_dir, model_rhs, int.vec=1:3)$solution
##  [1] 1 4 0
#
# You can get sensitivities in the integer case, but they're harder to
# interpret.
#
lp ("max", model_obj, model_con, model_dir, model_rhs, int.vec=1:3, compute.sens=TRUE)$duals
##  [1] 1 0 0 7 0
#
```




### Quadratic Programming by quadprog


solving quadratic programming problems of the form min(-d^T b + 1/2 b^T D b) with the constraints A^T b >= b_0.


$$\begin{align*} 
    \text{minimize: }   - 5 x_2 + 1/2 (x_1^2 + x_2^2 + x_3^2)\\
\text{subject to: }     -4 x_1 - 3 x_2       & >= -8 \\
                         2 x_1 +   x_2       & >= 2 \\
                         - 2 x_2 + x_3 & >= 0

\end{align*}$$



D is diagnal matix 

```{r}
library(quadprog)

## Assume we want to minimize: -(0 5 0) %*% b + 1/2 b^T D b
## under the constraints:      A^T b >= b0
##          1 0 0
##      D = 0 1 0
##          0 0 1
## with b0 = (-8,2,0)^T
## and      (-4  2  0) 
##      A = (-3  1 -2)
##          ( 0  0  1)
## we can use solve.QP as follows:
##
Dmat       <- matrix(0,3,3)
diag(Dmat) <- 1
dvec       <- c(0,5,0)
Amat       <- matrix(c(-4,-3,0,2,1,0,0,-2,1),3,3)
bvec       <- c(-8,2,0)
solve.QP(Dmat,dvec,Amat,bvec=bvec)

```





### Football MIP by glpk



trying to pick the best possible fantasy football team given different constraints.  goal is to pick the players that maximize the sum of their projected points.

The constraints are:

1) The team must include:

    - 1 QB
  
    - 2 RBs

    - 2 WRs

    - 1 TE

2) A player's risk must not exceed 6

3) The sum of the players' costs must not exceed 300. 


```{r}
#http://stackoverflow.com/questions/15147398/optimize-value-with-linear-or-non-linear-constraints-in-r



# We are going to solve:
# maximize f'x subject to A*x <dir> b
# where:
#   x is the variable to solve for: a vector of 0 or 1:
#     1 when the player is selected, 0 otherwise,
#   f is your objective vector,
#   A is a matrix, b a vector, and <dir> a vector of "<=", "==", or ">=",
#   defining your linear constraints.


name <- c("Aaron Rodgers","Tom Brady","Arian Foster","Ray Rice","LeSean McCoy","Calvin Johnson","Larry Fitzgerald","Wes Welker","Rob Gronkowski","Jimmy Graham")


# position
pos <- c("QB","QB","RB","RB","RB","WR","WR","WR","TE","TE")

# points
pts <- c(167, 136, 195, 174, 144, 135, 89, 81, 114, 111) 

risk <- c(2.9, 3.4, 0.7, 1.1, 3.5, 5.0, 6.7, 4.7, 3.7, 8.8) 

cost <- c(60, 47, 63, 62, 40, 60, 50, 35, 40, 40) 

#mydata <- data.frame(name, pos, pts, risk, cost) 

# number of variables
num.players <- length(name)



# objective:
f <- pts



# the variable are booleans
var.types <- rep("B", num.players)



# the constraints
A <- rbind(as.numeric(pos == "QB"), # num QB
           as.numeric(pos == "RB"), # num RB
           as.numeric(pos == "WR"), # num WR
           as.numeric(pos == "TE"), # num TE
           diag(risk),              # player's risk, a set of multiple constraints
           cost)                    # total cost

# diretion
dir <- c("==",
         "==",
         "==",
         "==",
         rep("<=", num.players),
         "<=")

# rhs b vector
b <- c(1,
       2,
       2,
       1,
       rep(6, num.players),
       300)



# solve
library(Rglpk)
sol <- Rglpk_solve_LP(obj = f, mat = A, dir = dir, rhs = b,
                      types = var.types, max = TRUE)
sol
# $optimum
# [1] 836                      ### <- the optimal total points



# $solution
#  [1] 1 0 1 0 1 1 0 1 1 0     ### <- a `1` for the selected players

# $status
# [1] 0                        ### <- an optimal solution has been found

# your dream team
name[sol$solution == 1]
# [1] "Aaron Rodgers"  "Arian Foster"   "LeSean McCoy"
# [4] "Calvin Johnson" "Wes Welker"     "Rob Gronkowski


# total cost
sum(cost * sol$solution)

```





